FHJ: A Formal Model for Resolving Unintended Conflicts in Multiple Inheritance

### 2017-06-14 Aplas17/sections/Formalization.tex
A quick review on related work.

\subsection{Properties}
Previously the definitions of our model are given, now we should proceed to prove the type soundness of 
the model, which relates typing to computation. The type soundness states that, if an expression is 
well-typed, then after many reduction steps it must reduce to a value, with its annotation to be subtype of the original expression type.
Following the FeatherweightJava paper~\cite{Igarashi01FJ}, the type-soundness theorem 
(Theorem~\ref{theorem_soundness}) is proved by using the standard technique of subject reduction (Theorem~\ref{theorem_subject})
and progress (Theorem~\ref{theorem_progress})~\cite{Wright1994}. In Theorem~\ref{theorem_progress} ``$\#(\overline{x})$'' denotes the number of
elements.

>>>>

\subsection{Properties}
Previously the definitions of our model are given, now we should proceed to prove the type soundness of 
the model, which relates typing to computation. The type soundness states that, if an expression is 
well-typed, then after many reduction steps it must reduce to a value, with its annotation to be subtype of the original expression type.
Following the Featherweight Java paper~\cite{Igarashi01FJ}, the type-soundness theorem 
(Theorem~\ref{theorem_soundness}) is proved by using the standard technique of subject reduction (Theorem~\ref{theorem_subject})
and progress (Theorem~\ref{theorem_progress})~\cite{Wright1994}. In Theorem~\ref{theorem_progress} ``$\#(\overline{x})$'' denotes the number of
elements.

### 2017-06-14 Aplas17/sections/RelatedWork.tex
A quick review on related work.

The above mentioned models/languages support multiple inheritance, and they handle method confliction in the same way: requiring 
programmers to explicitly resolve ambiguity and disallowing two methods with the same signature from two different libraries 
to co-exit. This is the common drawback of the above mentioned models versus our model.

\subsection{Resolving Unintended Method Confliction}
There are still a few languages that already realized the problem and explored a bit. We discuss them one by one.

\noindent {\bf C++ model.}
As discussed in Section~\ref{sec:overview}, C++ supports very flexible inheritance mode and allows programmers to choose different 
confliction resolution approaches via normal or virtual inheritance. Generally speaking, C++ model is much more complex 
and flexible than our approach. For example, given the following code
\begin{lstlisting}[language=Java]
class A { public: void m() {cout << "MA" << endl;}};
class B { public: void m() {cout << "MB" << endl;}};
class C : public A, public B { 
	void m() {cout << "MC" << endl;}
};
void func(A* a) { a->m(); }
int main() {
	C* c = new C();
	c->B::m();
	func(c); 
	return 0; //Running result: MB MA
}
\end{lstlisting}
The running result is $MB \; MA$, meaning that it uses static dispatch (looks at the static types) when doing method lookup. 
On calling \texttt{func(c)}, no matter the dynamic type of \texttt{c} is class \texttt{C}, the method call still dispatches to 
the method \texttt{m} in class \texttt{A} because of static dispatch.
However, we can alter the code a little bit with \textbf{\texttt{virtual}} methods and the result will be totally different:
\begin{lstlisting}[language=c++]
class A { 
	public: virtual void m() {cout << "MA" << endl;}
};
class B { 
	public: virtual void m() {cout << "MB" << endl;}
};
class C : public A, public B { 
    public: virtual void m() {cout << "MC" << endl;}
};
void func(A* a) { a->m(); }
int main() {
	C* c = new C();
	c->B::m();
	func(c); 
	return 0; //Running result: MB MC
}
\end{lstlisting}
Now the running result will be $MB \; MC$. With virtual method, dynamic dispatch is used and 
method lookup algorithm will find the most specific method definition of $m$, which is the definition 
in class $C$. Although C++ support this flexibility, the drawback compared to our model is
obvious: it does not support updating both $A.m$ and $B.m$ in class $C$.

@@

\noindent  {\bf \csharp Explicit method implementation.}
Explicit method implementation is a special feature supported by \csharp. As described in \csharp
documentation~\cite{csharpdoc}, a class that implements an interface can explicitly implement a member of that
interface. When a member is explicitly implemented, it can only be accessed through an instance
of the interface. Explicit interface implementation allows the programmer to inherit two interfaces 
that share the same member names and give each interface member a separate implementation. 
Explicit interface member implementations have two advantages:
Explicit interface member implementations allow interface implementations to be excluded 
from the public interface of a class. This is particularly useful when a class implements an internal 
interface that is of no interest to a consumer of that class or struct.
Explicit interface member implementations allow disambiguation of interface members with the 
same signature. This is the the advantage that is similar to ours' model. However, there are two 
diffirencies: firstly, method (default) implementations are not allowed in \csharp interfaces; 
secondly, when handling unintended method confliction, multi-way method updating is not supported 
in \csharp.

@@

\subsection{Hirarchical Dispatch}
As we have discussed before, the mix of static and dynamic dispatch is needed by the literature 
pay little attention to this. In the prototype-based language \self~\cite{Chambers1991}, inheritance is a basic feature.
It does not include classes but instead allow individual objects to inherit from (or delegate to) other objects. 
Although it is different than class-based languages, the multiple inheritance model is somehow similar. The \self 
language support multiple (object) inheritance in a clever way. Not only the new inheritance
relation \emph{prioritized parents} is developed, but also the method lookup rule called 
\emph{sender path tiebreaker rule}, which is similar with our approach, using hirarchical 
information in method resolution, but in a prototype-based language setting.

\subsection{Formalization Based on FeatherweightJava}
FeatherweightJava~\cite{Igarashi01FJ} is a minimal core calculus of Java language, proposed by Igarashi et. al. Many models are based on 
FeatherweightJava, for example FeatherTrait~\cite{Liquori08ftj}, Featherweight defenders~\cite{goetz12fdefenders}, Jx~\cite{Nystrom2004}, as well as our model.
It provides the standard model of formalizing a Java-like object-oriented language and 
people can easily extends it to form a new language. In terms of formalization, they key novelty of our model is the use of 
static types as annotations along with various terms. As far as we know, this technique has not appear in literature before.

>>>>

The above mentioned models/languages support multiple inheritance, and they handle method conflicts in the same way, by
simply disallowing two methods with the same signature from two different units to coexist. The ambiguity is resolved either
by programmers explicitly or by setting a linearised composition. This is the common drawback of the above mentioned models versus our model.

\subsection{Resolving Unintended Method Conflicts}
There are still a few languages that partly realize the problem. They provide limited support for unintended conflicting methods.

\noindent {\bf C++ model.}
C++ supports very flexible inheritance models and allows programmers to choose either static dispatch or dynamic dispatch for method lookup.
It allows unintended conflicts and uses static dispatch to resolve them, as discussed in Section~\ref{sec:overview}. For example, given the following code
\begin{lstlisting}[language=Java]
class A { public: void m() {cout << "MA" << endl;} };
class B { public: void m() {cout << "MB" << endl;} };
class C : public A, public B { 
	void m() {cout << "MC" << endl;}
};
void func(A* a) { a->m(); }
int main() {
	C* c = new C();
	c->B::m();
	func(c); 
	return 0; //Running result: MB MA
}
\end{lstlisting}
The running result is $MB \; MA$, meaning that it uses static dispatch (looks at the static types) for method \lstinline|m()|. 
On calling \texttt{func(c)}, in spite that the dynamic type of \texttt{c} is class \texttt{C}, the method call still dispatches to 
\lstinline|A.m|.
However, we can alter the code a little bit with \textbf{\texttt{virtual}} methods and the result will be totally different:
\begin{lstlisting}[language=c++]
class A { public: virtual void m() {cout << "MA" << endl;} };
class B { public: virtual void m() {cout << "MB" << endl;} };
class C : public A, public B { 
    public: virtual void m() {cout << "MC" << endl;}
};
void func(A* a) { a->m(); }
int main() {
	C* c = new C();
	c->B::m();
	func(c); 
	return 0; //Running result: MB MC
}
\end{lstlisting}
Now the running result will be $MB \; MC$. With virtual methods, dynamic dispatch is used and 
method lookup algorithm will find the most specific method definition of $m$, namely \lstinline|C.m| at this time.
Although C++ support this flexibility, unfortunately, dynamic dispatch on conflicting methods will be problematic.
Furthermore, it does not support partial overrides compared to our model.

@@

\noindent  {\bf \csharp{} Explicit method implementation.}
Explicit method implementation is a special feature supported by \csharp{}. As described in \csharp{}
documentation~\cite{csharpdoc}, a class that implements an interface can explicitly implement a member of that
interface. When a member is explicitly implemented, it can only be accessed through an instance
of the interface. Explicit interface implementation allows the programmer to inherit two interfaces 
that share the same member names and give each interface member a separate implementation.

Explicit interface member implementations have two advantages.
Firstly, they allow interface implementations to be excluded 
from the public interface of a class. This is particularly useful when a class implements an internal 
interface that is of no interest to a consumer of that class or struct.
Secondly, they allow disambiguation of interface members with the 
same signature. However, there are two critical differences to \MIM{}:
(1) default implementations are not allowed in \csharp{} interfaces; 
(2) there are only one-level partial overrides (implementations).

@@

\subsection{Hierarchical Dispatch}
As we have discussed before, the mix of static and dynamic dispatch is particularly useful under certain circumstances,
yet there is little attention from related work. In the prototype-based language \self~\cite{Chambers1991}, inheritance is a basic feature.
It does not include classes but instead allow individual objects to inherit from (or delegate to) other objects. 
Although it is different from class-based languages, the multiple inheritance model is somehow similar. The \self{}
language supports multiple (object) inheritance in a clever way. It not only develops the new inheritance
relation \emph{prioritized parents}, but also adopts \emph{sender path tiebreaker rule} for method lookup, which is similar with our approach, using hierarchical information in method resolution, but in a prototype-based language setting.\haoyuan{Do they have updates?}

\subsection{Formalization Based on Featherweight Java}
Featherweight Java~\cite{Igarashi01FJ} is a minimal core calculus of Java language, proposed by Igarashi et. al\haoyuan{Need cite}. There are
many models built on
Featherweight Java, including FeatherTrait~\cite{Liquori08ftj}, Featherweight defenders~\cite{goetz12fdefenders}, Jx~\cite{Nystrom2004}, and so on.
It provides the standard model of formalizing a Java-like object-oriented language, and 
is easily extensible. In terms of formalization, they key novelty of our model is the use of 
static types as annotations along with various terms. As far as we know, this technique has not appear in literature before.
